### 2024-06-15 out/01. Introduction/03. Main result.tex
latex only

Assume the maximum-area monotone sofa \(S\) with rotation angle \(\omega \in (0, \pi/2]\). Then we can assume \(S = \mathcal{M}(S)\) in \Cref{eqn:monotone} as taking further monotonization does not increase the area.\footnote{In fact, that \(S\) attains the maximum-area only guarantees that the gain \(\mathcal{M}(S) \setminus S\) of monotonization is of measure zero. We show in \Cref{thm:monotonization-fixpoint} that for any monotone sofa, we have \(\mathcal{M}(S) = S\).} For each angle \(t \in [0, \omega]\), let \(R_t : \mathbb{R}^2 \to \mathbb{R}^2\) be the rotation of \(\mathbb{R}^2\) around the origin, and let \(\mathbf{x}_S(t)\) be the coordinate of the inner corner of the rotated hallway \(L_t\). Then the hallway \(L_t = R_t(L) + \mathbf{x}_S(t)\) rotated counterclockwise by angle \(t \in [0, \pi/2]\) is determined by its inner corner \(\mathbf{x}_S(t)\). So the parametrization \(\mathbf{x}_S : [0, \omega] \to \mathbb{R}^2\) of the inner corner with respect to angle \(t\), defined as the \emph{rotation path} of \(S\) in \autocite{romikDifferentialEquationsExact2018}, determines the monotone sofa
\[
S = H \cap V_\omega \cap \bigcap_{0 \leq t \leq \omega} \left( R_t(L) + \mathbf{x}_S(t) \right) 
\]
completely. In particular, Romik in \autocite{romikDifferentialEquationsExact2018} derived Gerver’s sofa \(S = S_G\) by solving for the local optimality of area of \(S\) as a set of ordinary differential equations on \(\mathbf{x}_S\).{[}\^{}gerver-derivation{]}

@@

In particular, Gerver’s sofa \(S_G\) is a monotone sofa satisfying the injectivity condition (note that the blue trajectory of \(\mathbf{x}_{S_G}\) in \Cref{fig:gerver} is injective and never below the bottom line \(y=0\)). For any monotone sofa satisfying the injectivity condition, the following upper bound is established.

@@

So the upper bound of \Cref{thm:main} is effective on a domain of monotone sofas containing the conjectured optimum \(S_G\). The upper bound \(1 + \omega^2/2\) of \Cref{thm:main} maximizes at \(\omega = \pi/2\) with the value \(1 + \pi^2/8 = 2.2337\dots\), which is much closer to the lower bound \(2.2195\dots\) of Gerver than the currently best upper bound of \(2.37\) of Kallus and Romik (\Cref{eqn:best-bounds}). We conjecture that a monotone sofa of maximum area should satisfy the premise of \Cref{thm:main}.

@@

With \Cref{thm:main}, proving \Cref{con:injectivity} will imply the unconditional upper bound \(\mu_{\max} \leq 1 + \pi^2/2 = 2.2337\dots\).

The main idea for proving \Cref{thm:main} is to overestimate the area of a monotone sofa \(S\). For the sake of explanation, fix the rotation angle \(\omega = \pi/2\). Observe that the lower boundary of Gerver’s sofa \(S_G\) consists of two red ‘tails’ and one blue ‘core’ in \Cref{fig:gerver}. The core is parametrized by the inner corner \(\mathbf{x}_{S_G}(t)\) for the interval \(t \in [\varphi, \pi/2 - \varphi]\) with constant \(\varphi = 0.0392\dots\) \autocite{romikDifferentialEquationsExact2018}, and forms the majority of the lower boundary. The region below the two tails, trimmed out by the inner left and right walls of \(L_t\) respectively, constitutes only about \(0.28 \%\) of the area \(2.2195\dots\) of the whole sofa. Motivated by this, we define the overestimation \(\mathcal{A}_1\) of the area \(\mathcal{A}\) of a monotone sofa \(S\) with rotation angle \(\pi/2\) as: the area of the convex hull \(K\) of \(S\), that we call the \emph{cap} of \(S\), subtracted by the region enclosed by \(\mathbf{x}_S : [0, \omega] \to \mathbb{R}^2\) and the line \(y=0\).

Another key idea for proving \Cref{thm:main} is to consider the area of \(S\) as a function of the cap \(K\). By doing so, we turn the problem to a quadratic optimization problem in convex analysis. In particular, the overestimated area \(\mathcal{A}_1(K)\) as a function of the cap \(K\) turns out to be a \emph{concave} and \emph{quadratic} functional with respect to the cap \(K\). To justify this, we first construct the space \(\mathcal{K}_\omega\) of all possible caps \(K\) with a fixed rotation angle \(\omega\). The space \(\mathcal{K}_\omega\) itself is a convex space equipped with convex combination \(c_\lambda(K_1, K_2) = (1 - \lambda) K_1 + \lambda K_2\) defined by the Minkowski sum of convex bodies. Then the Brunn-Minkowski theory \cite{schneider_2013} is used to establish that \(\mathcal{A}_1\) is a quadratic functional on \(\mathcal{K}_\omega\). We use Mamikon’s theorem \autocite{mnatsakanianAnnularRingsEqual1997}, a theorem in classical geometry, to prove that \(\mathcal{A}_1\) concave on \(\mathcal{K}_\omega\). Finally, we introduce a calculus of variation based on the Brunn-Minkowski theory to find a local optimum \(K_1\) of \(\mathcal{A}_1\). Since \(\mathcal{A}_1\) is concave, the local optimum \(K_1\) is also a global optimum of \(\mathcal{A}_1\), and \Cref{thm:main} is established by computing the maximum value \(\mathcal{A}_1(K_1) = 1 + \omega^2/2\) of \(\mathcal{A}_1\).

For the rotation angle \(\omega = \pi/2\), the maximizer of \(\mathcal{A}_1\) gives an unmovable sofa \(S_1\) of area \(1 + \pi^2/8 = 2.2337\dots\) and width \(\pi\) very close to the area of Gerver’s sofa \(S_G\) (see \Cref{fig:presofa}). The shape of \(S_1\) is very close to \(S_G\), and cutting away the region under the red curves from \(S_1\) gives a valid moving sofa of area approximately \(2.2009\dots\), which is again very close to \(S_G\) in its shape and area. The boundary of \(S_1\) consists of three line segments and three parametric curves. The right side of \(S_1\) is parametrized by the curve \(\gamma : [0, \pi/2] \to \mathbb{R}^2\) with \(\gamma(0) = (1, 1)\) and \(\gamma'(t) = t(\cos t, -\sin t)\), which ends at \(\gamma(\pi/2) = (\pi/2, 0)\). The left side of \(S_1\) is parametrized by the curve symmetric to \(\gamma\) along the \(y\)-axis. The bottom middle part of \(S_1\) is parametrized by the curve \(\mathbf{x}_{S_1} : [0, \pi/2] \to \mathbb{R}^2\) with \(\mathbf{x}_{S_1}(0) = (\pi/2-1, 0)\) and
\[
\mathbf{x}_{S_1}'(t) = -t (\cos t, \sin t) + (\pi/2- t) (-\sin t, \cos t)
\]
which is symmetric along the \(y\)-axis. The endpoints of the three curves are connected by three horizontal line segments of length 1 or 2 to form the boundary of \(S_1\).

\begin{figure}
\centering
\includesvg[width=1\textwidth,height=\textheight]{images/presofa-combined.svg}
\caption{The maximizing shape \(S_1\) of \(\mathcal{A}_1\) of area \(1 + \pi^2/8 = 2.2337\dots\). The regions below two tails (red curves) stick out of the hallway \(L\) during the movement of \(S_1\) in \(L\) (left). The shape \(S_1\) is very similar to Gerver’s sofa \(S_G\) whose boundary is drawn in dotted lines (right).}
\label{fig:presofa}
\end{figure}

>>>>

Assume the maximum-area monotone sofa \(S\) with rotation angle \(\omega \in (0, \pi/2]\). Then we can assume \(S = \mathcal{M}(S)\) in \Cref{eqn:monotone} as taking further monotonization does not increase the area.\footnote{In fact, that \(S\) attains the maximum-area only guarantees that the gain \(\mathcal{M}(S) \setminus S\) of monotonization is of measure zero. We show in \Cref{thm:monotonization-fixpoint} that for any monotone sofa, we have \(\mathcal{M}(S) = S\).} For each angle \(t \in [0, \omega]\), let \(R_t : \mathbb{R}^2 \to \mathbb{R}^2\) be the rotation of \(\mathbb{R}^2\) around the origin, and let \(\mathbf{x}_S(t)\) be the coordinate of the inner corner of the rotated hallway \(L_t\). Then the hallway \(L_t = R_t(L) + \mathbf{x}_S(t)\) rotated counterclockwise by angle \(t \in [0, \pi/2]\) is determined by its inner corner \(\mathbf{x}_S(t)\). So the parametrization \(\mathbf{x}_S : [0, \omega] \to \mathbb{R}^2\) of the inner corner with respect to angle \(t\), defined as the \emph{rotation path} of \(S\) in \autocite{romikDifferentialEquationsExact2018}, determines the monotone sofa
\[
S = H \cap V_\omega \cap \bigcap_{0 \leq t \leq \omega} \left( R_t(L) + \mathbf{x}_S(t) \right) 
\]
completely. In particular, Romik in \autocite{romikDifferentialEquationsExact2018} derived Gerver’s sofa \(S = S_G\) by solving for the local optimality of area of \(S\) as a set of ordinary differential equations on \(\mathbf{x}_S\).

@@

In particular, Gerver’s sofa \(S_G\) is a monotone sofa satisfying the injectivity condition. Note that the blue trajectory of \(\mathbf{x}_{S_G}\) in \Cref{fig:gerver} is injective and never below the bottom line \(y=0\). For any monotone sofa satisfying the injectivity condition, the following upper bound is established.

@@

The upper bound \(1 + \omega^2/2\) of \Cref{thm:main} maximizes at \(\omega = \pi/2\) with the value \(1 + \pi^2/8 = 2.2337\dots\), which is much closer to the lower bound \(2.2195\dots\) of Gerver than the currently best upper bound of \(2.37\) of Kallus and Romik (\Cref{eqn:best-bounds}). We conjecture that a monotone sofa of maximum area should satisfy the premise of \Cref{thm:main}.

@@

With \Cref{thm:main}, proving \Cref{con:injectivity} would imply the unconditional upper bound \(\mu_{\max} \leq 1 + \pi^2/2 = 2.2337\dots\). Since Gerver’s sofa \(S_G\) satisfies the injectivity condition, \Cref{con:injectivity} is a weakening of Gerver’s conjecture \(\mu_{\max} = \mu_G\).

The main idea for proving \Cref{thm:main} is to overestimate the area of a monotone sofa \(S\) (see \Cref{fig:a1-upper-bound}). For the sake of explanation, fix the rotation angle \(\omega = \pi/2\). Observe that the lower boundary of Gerver’s sofa \(S_G\) consists of two red ‘tails’ and one blue ‘core’ in \Cref{fig:gerver}. The core is parametrized by the inner corner \(\mathbf{x}_{S_G}(t)\) for the interval \(t \in [\varphi, \pi/2 - \varphi]\) with constant \(\varphi = 0.0392\dots\) \autocite{romikDifferentialEquationsExact2018}, and forms the majority of the lower boundary. The region below the two tails, trimmed out by the inner left and right walls of \(L_t\) respectively, constitutes only about \(0.28 \%\) of the area \(2.2195\dots\) of the whole sofa. Motivated by this, we define the overestimation \(\mathcal{A}_1\) of the area of a monotone sofa \(S\) with rotation angle \(\pi/2\) as: the area of the convex hull \(K\) of \(S\), that we call the \emph{cap} of \(S\), subtracted by the region enclosed by \(\mathbf{x}_S : [0, \omega] \to \mathbb{R}^2\) and the line \(y=0\).

\begin{figure}
\centering
\includesvg[width=0.7\textwidth,height=\textheight]{images/a1-upper-bound.svg}
\caption{Overestimation \(\mathcal{A}_1(K)\) of the area of a monotone sofa with cap \(K\), including the two red ‘tails’ but excluding the blue ‘core’.}
\label{fig:a1-upper-bound}
\end{figure}

Another key idea for proving \Cref{thm:main} is to consider the area of \(S\) as a function of the cap \(K\). The overestimated area \(\mathcal{A}_1(K)\) in \Cref{fig:a1-upper-bound}, as a function of the cap \(K\), turns out to be \emph{concave} and \emph{quadratic} with respect to \(K\). The space \(\mathcal{K}_\omega\) of all caps \(K\) with a fixed rotation angle \(\omega\) forms a convex space equipped with the convex combination \(c_\lambda(K_1, K_2) = (1 - \lambda) K_1 + \lambda K_2\) defined by the Minkowski sum of convex bodies. Then the Brunn-Minkowski theory on convex bodies \cite{schneider_2013} is used to establish that \(\mathcal{A}_1\) is a quadratic functional on \(\mathcal{K}_\omega\). We use Mamikon’s theorem \autocite{mnatsakanianAnnularRingsEqual1997}, a theorem in classical geometry, to prove that \(\mathcal{A}_1\) concave on \(\mathcal{K}_\omega\) (\Cref{fig:mamikon-sofa}).\footnote{For our application, we rigorously state and prove a generalization (\Cref{thm:mamikon}) of Mamikon’s theorem that is effective on any planar convex body with potentially non-differentiable boundary, which could be of independent interest.} Finally, we introduce a calculus of variation based on the Brunn-Minkowski theory to find a global optimum \(K_1\) of \(\mathcal{A}_1\). \Cref{thm:main} is established by computing the maximum value \(\mathcal{A}_1(K_1) = 1 + \omega^2/2\) of \(\mathcal{A}_1\).

For the rotation angle \(\omega = \pi/2\), the maximizer of \(\mathcal{A}_1\) gives an unmovable sofa \(S_1\) of area \(1 + \pi^2/8 = 2.2337\dots\) and width \(\pi\) very close to the area of Gerver’s sofa \(S_G\) (see the right side of \Cref{fig:presofa}). The shape of \(S_1\) is very close to \(S_G\), and cutting away the region under the red curves from \(S_1\) gives a valid moving sofa of area approximately \(2.2009\dots\), which is again very close to \(S_G\) in its shape and area. The boundary of \(S_1\) consists of three line segments and three parametric curves. The right side of \(S_1\) is parametrized by the curve \(\gamma : [0, \pi/2] \to \mathbb{R}^2\) with \(\gamma(0) = (1, 1)\) and \(\gamma'(t) = t(\cos t, -\sin t)\), which ends at \(\gamma(\pi/2) = (\pi/2, 0)\). The left side of \(S_1\) is parametrized by the curve symmetric to \(\gamma\) along the \(y\)-axis. The bottom middle part of \(S_1\) is parametrized by the curve \(\mathbf{x}_{S_1} : [0, \pi/2] \to \mathbb{R}^2\) with \(\mathbf{x}_{S_1}(0) = (\pi/2-1, 0)\) and
\[
\mathbf{x}_{S_1}'(t) = -t (\cos t, \sin t) + (\pi/2- t) (-\sin t, \cos t)
\]
which is symmetric along the \(y\)-axis. The endpoints of the three curves are connected by three horizontal line segments of length 1 or 2 to form the boundary of \(S_1\).

### 2024-06-15 out/01. Introduction/04. Overview of Sections.tex
latex only

\Cref{sec:monotone-sofas} proves \Cref{thm:monotonization-is-sofa} that any moving sofa can be enlarged to a monotone sofa, and proves structural \Cref{thm:monotone-sofa-structure} and \Cref{thm:niche-in-cap} on monotone sofas. Using this, \Cref{sec:sofa-area-functional-a} reduces the moving sofa problem to the maximization of the \emph{sofa area functional} \(\mathcal{A} : \mathcal{K}_\omega \to \mathbb{R}\) on the space of caps \(\mathcal{K}_\omega\). \Cref{sec:conditional-upper-bound-a1} proves the main \Cref{thm:main} by establishing the upper bound \(1 + \omega^2/2\) of the sofa area \(\mathcal{A}\) using the upper bound \(\mathcal{A}_1\). Each chapter starts with an overview of the chapter.

\Cref{sec:convex-bodies} proves numerous properties of an arbitrary planar convex body \(K\) that
we will use thoroughly in this paper. So the logical ordering of this paper is
\Cref{sec:notations-and-conventions}, followed by \Cref{sec:convex-bodies}, then the chapters
starting \Cref{sec:monotone-sofas}. A logically inclined reader may read in this ordering to verify
the correctness of all arguments. On the other hand, readers who are interested in the overall idea
may either start by reading the chapters in order and refer to the appendix when needed, or skim the introduction of each section of \Cref{sec:convex-bodies} and then read the rest of the paper.

>>>>

\Cref{sec:monotone-sofas} proves \Cref{thm:monotonization-is-sofa} that any moving sofa can be enlarged to a monotone sofa, and proves structural \Cref{thm:monotone-sofa-structure} and \Cref{thm:niche-in-cap} on monotone sofas. Using this, \Cref{sec:sofa-area-functional-a} reduces the moving sofa problem to the maximization of the \emph{sofa area functional} \(\mathcal{A} : \mathcal{K}_\omega \to \mathbb{R}\) on the space of caps \(\mathcal{K}_\omega\). \Cref{sec:conditional-upper-bound-a1} proves the main \Cref{thm:main} by establishing the upper bound \(1 + \omega^2/2\) of the sofa area \(\mathcal{A}\) using the upper bound \(\mathcal{A}_1\). Each section starts with an overview of its own.

\Cref{sec:convex-bodies} proves numerous properties of an arbitrary planar convex body \(K\) that we will use thoroughly in this paper. So the logical ordering of this paper is \Cref{sec:notations-and-conventions}, followed by \Cref{sec:convex-bodies}, then the sections starting \Cref{sec:monotone-sofas}. A logically inclined reader may read in this ordering to verify the correctness of all arguments. On the other hand, readers who are interested in the overall idea may either start by reading the sections in order and refer to the appendix when needed, or skim the introduction of each section of \Cref{sec:convex-bodies} and then read the rest of the paper.

### 2024-06-15 out/10. Notations and conventions/03. Convex body.tex
latex only

For reference, the notion of cap \(K\) as a kind of convex bodies is in \Cref{def:cap}. The space of all caps \(\mathcal{K}_\omega\) with rotation angle \(\omega\) is in \Cref{def:cap-space}. The vertices \(A_K^{\pm}(t)\), \(C_K^{\pm}(t)\) of a cap \(K\) is in \Cref{def:cap-vertices}. The upper boundary \(\delta K\) of a cap \(K\) is in \Cref{def:upper-boundary-of-cap}.

>>>>

For reference, the notion of cap \(K\) as a kind of convex bodies is defined in \Cref{def:cap}. The space of all caps \(\mathcal{K}_\omega\) with rotation angle \(\omega\) is defined in \Cref{def:cap-space}. The vertices \(A_K^{\pm}(t)\), \(C_K^{\pm}(t)\) of a cap \(K\) is defined in \Cref{def:cap-vertices}. The upper boundary \(\delta K\) of a cap \(K\) is defined in \Cref{def:upper-boundary-of-cap}.

### 2024-06-15 out/10. Notations and conventions/04. Area functionals.tex
latex only

The integral in \Cref{def:curve-area-functional} is the Lebesgue-Stieltjes integral, for which we again refer to \autocite{steinRealAnalysisMeasure2005}. By change of variables (e.g.~Equation 2 of \autocite{falknerSubstitutionRuleLebesgue2012}), the value of \(\mathcal{I}(\mathbf{x}) = \mathcal{I}(\mathbf{x} \circ \alpha)\) is the same even if we replace \(\mathbf{x}\) with a reparametrization \(\mathbf{x} \circ \alpha : [a', b'] \to \mathbb{R}^2\) of curve \(\Gamma\), where \(\alpha : [a', b'] \to [a, b]\) is any monotonically increasing, continuous, and surjective function. In particular, for any parametrization \(\mathbf{x}\) of the line segment from point \(p\) to \(q\), the value \(\mathcal{I}(\mathbf{x})\) is equal to \(1/2 \cdot (p \times q)\).

@@

For reference, the function \(\mathcal{A}(K)\) is the \emph{sofa area functional} on cap \(K\) in \Cref{def:sofa-area-functional}. \(\mathcal{A}_1(K)\) is a conditional upper bound of \(\mathcal{A}(K)\) in \Cref{def:a1}.

>>>>

The integral in \Cref{def:curve-area-functional} is the Lebesgue-Stieltjes integral, for which we again refer to \autocite{steinRealAnalysisMeasure2005}. By change of variables (e.g.~Equation 2 of \autocite{falknerSubstitutionRuleLebesgue2012}), the value of \(\mathcal{I}(\mathbf{x}) = \mathcal{I}(\mathbf{x} \circ \alpha)\) is the same even if we replace \(\mathbf{x}\) with a reparametrization \(\mathbf{x} \circ \alpha : [a', b'] \to \mathbb{R}^2\) of curve \(\Gamma\), where \(\alpha : [a', b'] \to [a, b]\) is any monotonically increasing, continuous, and surjective function. In particular, for any parametrization \(\mathbf{x}\) of the line segment from point \(p\) to \(q\), its curve area functional \(\mathcal{I}(\mathbf{x})\) is equal to \(1/2 \cdot (p \times q)\).

@@

For reference, the function \(\mathcal{A}(K)\) is the \emph{sofa area functional} on cap \(K\) defined in \Cref{def:sofa-area-functional}. \(\mathcal{A}_1(K)\) is a conditional upper bound of \(\mathcal{A}(K)\) defined in \Cref{def:a1}.

### 2024-06-15 out/10. Notations and conventions/05. Parts of hallway.tex
latex only

For reference, the \emph{tangent hallway} \(L_S(t)\) of a shape \(S\) will be defined in \Cref{def:tangent-hallway}. The corresponding parts \(\mathbf{x}_S(t), \mathbf{y}_S(t), a_S(t), b_S(t), c_S(t), d_S(t), Q^+_S(t), Q^-_S(t)\) of the tangent hallway will be defined in \Cref{def:rotating-hallway-parts}.

>>>>

For reference, the \emph{tangent hallway} \(L_S(t)\) of a shape \(S\) will be defined in \Cref{def:tangent-hallway}. The corresponding parts \(\mathbf{x}_S(t)\), \(\mathbf{y}_S(t)\), \(a_S(t)\), \(b_S(t)\), \(c_S(t)\), \(d_S(t)\), \(Q^+_S(t)\), \(Q^-_S(t)\) of the tangent hallway will be defined in \Cref{def:rotating-hallway-parts}.

### 2024-06-15 out/A1. Convex bodies/07. Lebesgue-Stieltjes measure.tex
latex only

Most literature constructs \(dF\) for any right-continuous and monotonically increasing \(F : \mathbb{R} \to \mathbb{R}\) \autocite{steinRealAnalysisMeasure2005,hewittRealAbstractAnalysis2013,halmos2013measure}.

>>>>

Most literature constructs \(dF\) for any right-continuous and monotonically increasing \(F : \mathbb{R} \to \mathbb{R}\) \autocite{steinRealAnalysisMeasure2005,hewittRealAbstractAnalysis1965,halmos2013measure}.

### 2024-06-15 out/A1. Convex bodies/15. Parametrization of boundary.tex
Title Case
latex only

\subsection{Definition of parametrization}

@@

\subsection{Theorems on parametrization}

@@

\subsection{Injectivity of parametrization}

@@

\begin{proof}
(of \Cref{thm:param-positive-jordan})

That \(\mathbf{b}_K^{t, t + 2\pi}\) is an arc-length parametrization of \(\partial K\) comes from \Cref{thm:param-segment} and \Cref{thm:boundary-is-union-all-edges}.

@@

\subsection{Closed interval}

>>>>

\subsubsection{Definition of Parametrization}

@@

\subsubsection{Theorems on Parametrization}

@@

\subsubsection{Injectivity of Parametrization}

@@

\begin{proof}[Proof of \Cref{thm:param-positive-jordan}]
That \(\mathbf{b}_K^{t, t + 2\pi}\) is an arc-length parametrization of \(\partial K\) comes from \Cref{thm:param-segment} and \Cref{thm:boundary-is-union-all-edges}.

@@

\subsubsection{Parametrization on Closed Interval}
